Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2016-11-10 apj.tex
more changes for Gerard and myself

\subsection{Model}
We assume 
that  peak intrinsic $UBVRI$ magnitudes are linearly dependent
on the
 equivalent widths of the CaII H\&K and SiII$\lambda$4141 spectral features
$EW_{Ca}$, $EW_{Si}$,
and $\lambda_{Si}$ the wavelength of the minimum of 
the SiII$\lambda6355$ feature:
these spectral features are associated with SN~Ia  spectroscopic diversity  
\citep{2006PASP..118..560B, 2008A&A...492..535A, 2009A&A...500L..17B, 2009PASP..121..238B, 2009ApJ...699L.139W, 2011ApJ...729...55F}.
The explicit omission of light-curve shape in our model is compensated by its proxy
$EW_{Si}$ at peak brightness
\citep{2008A&A...492..535A, 2011A&A...529L...4C}. 
Residual dispersion is described by a Normal distribution with a parameterized covariance matrix
$C_c$.  A grey magnitude offset, $\Delta$, is included for each supernova
to capture intrinsic dispersion and peculiar-velocity errors introduced when converting fluxes to luminosity.
The observed magnitudes are linearly dependent on the
extrinsic-color parameters $k_0$ and $k_1$.  
The observables
$U_o, B_o, V_o, R_o, I_o$, $EW_{Ca,o}$, $EW_{Si,o}$, $\lambda_{Si,o}$
shown in Table~\ref{data:tab} have Gaussian measurement uncertainty with covariance $C$.
Unlike the parameters associated with
the spectral measurements, $EW_{Ca}$, $EW_{Si}$ and $\lambda_{Si}$,  the latent
parameters $k_0$ and $k_1$ are not directly associated
with observables but are rather drawn from the set of peak magnitudes.
The model 
\color{red}
likelihood density
\color{black}
is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_0+\alpha_0 EW_{Ca} + \beta_0 EW_{Si} + \eta_0 \lambda_{Si} \\
c_1+\alpha_1 EW_{Ca} + \beta_1 EW_{Si} + \eta_1 \lambda_{Si}  \\
c_2+\alpha_2 EW_{Ca} + \beta_2 EW_{Si} + \eta_2 \lambda_{Si} \\
c_3+\alpha_3 EW_{Ca} + \beta_3 EW_{Si} + \eta_3 \lambda_{Si} \\
c_4+\alpha_4 EW_{Ca} + \beta_4 EW_{Si}+ \eta_4 \lambda_{Si}
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{1} k_0 +\gamma^1_{1} k_1 \\B +\gamma^0_{1} k_0 +\gamma^1_{1} k_1 \\
V+\gamma^0_{2} k_0+\gamma^1_{2} k_1\\R+\gamma^0_{3} k_0 + \gamma^1_{3} k_1\\I+\gamma^0_{4} k_0+\gamma^1_{4} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}
The global parameters $c$, $\alpha$, $\beta$,  and $\eta$,  are the intercepts and slopes of the linear relationships that
relate intrinsic magnitudes with the spectral-feature parameters.
(In this article we freely interchange the parameter indices with  $01234$ and $UBVRI$.)
The global parameters $\gamma^0$, $\gamma^1$  are the slopes that connect the extrinsic-color
parameters to observed magnitudes.

@@

@@

We find significant non-zero values for $\alpha$ and $\beta$, indicating that $EW_{Ca}$ and $EW_{Si}$ are indicators of broadband
peak magnitudes.
This validates our hypothesis that spectral indicators
are tracers of supernova absolute magnitude.  On the other hand, the values of $\eta$ (the coefficients attached to $\lambda_{Si}$) are consistent with zero
to  within one standard deviation.
The effect of spectral parameters on color is shown in the rows of $\alpha_X/\alpha_V-1$,  $\beta_X/\beta_V-1$, and  $\eta_X/\eta_V-1$
in Table~\ref{global:tab}:
values of zero signify no color changes associated with magnitude changes.
Both $EW_{Ca}$ and $EW_{Ca}$ are associated with color changes, though not in $B-V$ specifically.
We do not detect a significant association between
$\lambda_{Si}$ and color.

@@

The quantity $R^F_V = A^F_V/E^F(B-V)$ for an individual supernova is unspecified without knowledge of the intrinsic color.
For a qualitative view, we choose the zeros for $V$-band extinction and color-excess such that the bluest supernova roughly correspond to $E(B-V)=0$,
$A_V=0$, and such that the ridge line is close to constant.
With this ansatz, we
estimate $R^T_V$ per supernova shown in Figure~\ref{avebv:fig}, with the expectations of the
 \citetalias{1999PASP..111...63F} model for fixed $R^F_V$ overplotted.
With this choice of zeropoints, the ridge line corresponds to $R_V^F\lesssim 2.5$, and even lower values of $R_V^F$ below the ridge line.
We emphasize that the shape of the locus in this plot (though not the relative positions of the points)  and the flatness
of the ridge line is specific to our choice of zeropoints, and that our model does not specify the zeropoints.
A specification of per-supernova $R_V$ requires a model that can identify supernovae that suffer no extrinsic effects,
for example by modeling  a constraining distribution for the extrinsic parameters \citep{2016arXiv160904470M}.
\color{black}

@@

Until now we have considered $k_0$ and $k_1$ without association with any specific dust model, but they could be interpreted as
parameters of a single  dust model.
\color{red}
Plotted in Figure~\ref{avebv:fig} are the expectations for measured $A(V)$, $E(B-V)$, and $R^T_V$ for $R^F_V=1.1, 1.9, 2.5, 3.1$
within the context of the \citetalias{1999PASP..111...63F} model.
Note that for our broadband measurements, the relationship between $A(V)$ and $E(B-V)$ is subtly
non-linear, and $R^T_V$ is not constant for a fixed value of $R^F_V$.
\color{black}

The \citet{1989ApJ...345..245C} dust law can be expressed as a two-parameter linear model
\begin{equation}
A_X = a(X)  A_V + b(X) E(B-V).
\end{equation}
The \citet{1999PASP..111...63F} (hereafter referred to as \citetalias{1999PASP..111...63F}) model is approximately described by a linear model with
the above form, with
$a(U,B,V,R,I)=\{0.96,   1.00,   1.00,   0.97,   0.78\}$ and $b(U,B,V,R,I)=\{  1.64,   1.06,   0.16,  -0.40,  -0.52\}$ for the case of
$R^F_V=2.5$
\color{red}
and $A^F_V=0.1$,
for the measurement of the updated
\citet{2007ApJ...663.1187H} SN~Ia template at $B$-band peak as seen through our synthetic passbands.
\color{black}
This linear approximation
is good in the ranges of
\color{red}
 $R^F_V$ and $A^F_V$,
\color{black}
and wavelengths under consideration, in that the values of the components  vary by $<5$\%.

@@

\section{Supernova Model II: Two Extrinsic Parameters, One Intrinsic Parameter}
\label{model2:sec}
In \S\ref{model:sec} we show that the SNfactory data analyzed with an agnostic model with
two extrinsic parameters produce
magnitudes consistent
with two-parameter dust models that have been established for stars in the Milky Way, LMC, and SMC.
An extension to our model is now added to explore the possibility of
a possible additional
color parameter that could have its origins in supernova astrophysics.
The extended model is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_0+\alpha_0 EW_{Ca} + \beta_0 EW_{Si} + \eta_0 \lambda_{Si} + \delta_0 D\\
c_1+\alpha_1 EW_{Ca} + \beta_1 EW_{Si} + \eta_1 \lambda_{Si} + \delta_1 D \\
c_2+\alpha_2 EW_{Ca} + \beta_2 EW_{Si} + \eta_2 \lambda_{Si} + \delta_2 D\\
c_3+\alpha_3 EW_{Ca} + \beta_3 EW_{Si} + \eta_3 \lambda_{Si} + \delta_3 D\\
c_4+\alpha_4 EW_{Ca} + \beta_4 EW_{Si}+ \eta_4 \lambda_{Si} + \delta_4 D
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{1} k_0 +\gamma^1_{1} k_1 \\B +\gamma^0_{1} k_0 +\gamma^1_{1} k_1 \\
V+\gamma^0_{2} k_0+\gamma^1_{2} k_1\\R+\gamma^0_{3} k_0 + \gamma^1_{3} k_1\\I+\gamma^0_{4} k_0+\gamma^1_{4} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}
In all aspects it is identical to the previous model except for the addition of the term $\delta D$,
which affects the mean
absolute magnitudes:
each supernova has an intrinsic $D$ parameter that is linearly related to
absolute magnitudes through global coefficients $\delta$.  As with $k_0$ and $k_1$,  $D$ is a latent supernova parameter whose existence is inferred from
its effects on broadband magnitudes.

>>>>

\subsection{Model}
We assume 
that  peak intrinsic $UBVRI$ magnitudes are linearly dependent
on the
 equivalent widths of the CaII H\&K and SiII$\lambda$4141 spectral features
$EW_{Ca}$, $EW_{Si}$,
and $\lambda_{Si}$ the wavelength of the minimum of 
the SiII$\lambda6355$ feature:
these spectral features are associated with SN~Ia  spectroscopic diversity  
\citep{2006PASP..118..560B, 2008A&A...492..535A, 2009A&A...500L..17B, 2009PASP..121..238B, 2009ApJ...699L.139W, 2011ApJ...729...55F}.
The explicit omission of light-curve shape in our model is compensated by its proxy
$EW_{Si}$ at peak brightness
\citep{2008A&A...492..535A, 2011A&A...529L...4C}. 
Residual dispersion is described by a Normal distribution with a parameterized covariance matrix
$C_c$.  A grey magnitude offset, $\Delta$, is included for each supernova
to capture intrinsic dispersion and peculiar-velocity errors introduced when converting fluxes to luminosity.
The observed magnitudes are linearly dependent on the
extrinsic-color parameters $k_0$ and $k_1$.  
The observables
$U_o, B_o, V_o, R_o, I_o$, $EW_{Ca,o}$, $EW_{Si,o}$, $\lambda_{Si,o}$
shown in Table~\ref{data:tab} have Gaussian measurement uncertainty with covariance $C$.
Unlike the parameters associated with
the spectral measurements, $EW_{Ca}$, $EW_{Si}$ and $\lambda_{Si}$,  the latent
parameters $k_0$ and $k_1$ are not directly associated
with observables but are rather drawn from the set of peak magnitudes.
The model 
\color{red}
likelihood density
\color{black}
is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_0+\alpha_0 EW_{Ca} + \beta_0 EW_{Si} + \eta_0 \lambda_{Si} \\
c_1+\alpha_1 EW_{Ca} + \beta_1 EW_{Si} + \eta_1 \lambda_{Si}  \\
c_2+\alpha_2 EW_{Ca} + \beta_2 EW_{Si} + \eta_2 \lambda_{Si} \\
c_3+\alpha_3 EW_{Ca} + \beta_3 EW_{Si} + \eta_3 \lambda_{Si} \\
c_4+\alpha_4 EW_{Ca} + \beta_4 EW_{Si}+ \eta_4 \lambda_{Si}
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{1} k_0 +\gamma^1_{0} k_1 \\B +\gamma^0_{1} k_0 +\gamma^1_{1} k_1 \\
V+\gamma^0_{2} k_0+\gamma^1_{2} k_1\\R+\gamma^0_{3} k_0 + \gamma^1_{3} k_1\\I+\gamma^0_{4} k_0+\gamma^1_{4} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}
The global parameters $c$, $\alpha$, $\beta$,  and $\eta$,  are the intercepts and slopes of the linear relationships that
relate intrinsic magnitudes with the spectral-feature parameters.
(In this article we freely interchange the parameter indices with  $01234$ and $UBVRI$.)
The global parameters $\gamma^0$, $\gamma^1$  are the slopes that connect the extrinsic-color
parameters to observed magnitudes.

@@

\color{blue}
There remains degeneracies in the model.  One is illustrated by the  transformations $\Delta \rightarrow \Delta  + \epsilon_\Delta$,
 $\gamma^0 \rightarrow \gamma^0  + \epsilon_{\gamma^0}$, $\gamma^1 \rightarrow \gamma^1 + \epsilon_{\gamma^1}$under the condition
$$
\epsilon_\Delta  +  \epsilon_{\gamma^0} k_0+  \epsilon_{\gamma^1} k_1=0,
$$
i.e.\ shifts in the global coefficients $\gamma$ can be compensated by shifts in $\Delta$ for each supernova while maintaining 
$\langle \Delta \rangle=0$.
Another degeneracy is illustrated by $k_0 \rightarrow k_0 + (EW_{Ca}-\langle EW_{Ca}\rangle)\epsilon$,
$\alpha \rightarrow \alpha - \gamma \epsilon$, $c \rightarrow c + \langle EW_{Ca}\rangle \epsilon$, i.e. the
linear external terms can leak into the linear spectral feature corrections.
These degeneracies are not broken within the model, but rather are constrained by the priors, for which the
per-supernova parameters
$\Delta$, $EW$, $\lambda$, $k$ are uncorrelated.
\color{black}

@@

We find significant non-zero values for $\alpha$ and $\beta$, indicating that $EW_{Ca}$ and $EW_{Si}$ are indicators of broadband
peak magnitudes.
This validates our hypothesis that spectral indicators
are tracers of supernova absolute magnitude.  On the other hand, the values of $\eta$ (the coefficients attached to $\lambda_{Si}$) are consistent with zero
to  within one standard deviation.
The effect of spectral parameters on color is shown in the rows of $\alpha_X/\alpha_V-1$,  $\beta_X/\beta_V-1$, and  $\eta_X/\eta_V-1$
in Table~\ref{global:tab}:
values of zero signify no color changes associated with magnitude changes.
Both $EW_{Ca}$ and $EW_{Si}$ are associated with color changes, though not in $B-V$ specifically.
We do not detect a significant association between
$\lambda_{Si}$ and color.

@@

The quantity $R^F_V = A^F_V/E^F(B-V)$ for an individual supernova is unspecified without knowledge of the intrinsic color.
For a qualitative view, we choose the zeros for $V$-band extinction and color-excess such that the bluest supernova roughly correspond to $E(B-V)=0$,
$A_V=0$, and such that the ridge line is close to constant.
With this ansatz, we
estimate $R^T_V$ per supernova shown in Figure~\ref{avebv:fig}, with the expectations of the
 \citet{1999PASP..111...63F} (hereafter referred to as \citetalias{1999PASP..111...63F})  model for fixed $R^F_V$ overplotted.
With this choice of zeropoints, the ridge line corresponds to $R_V^F\lesssim 2.5$, and even lower values of $R_V^F$ below the ridge line.
We emphasize that the shape of the locus in this plot (though not the relative positions of the points)  and the flatness
of the ridge line is specific to our choice of zeropoints, and that our model does not specify the zeropoints.
A specification of per-supernova $R_V$ requires a model that can identify supernovae that suffer no extrinsic effects,
for example by modeling  a constraining distribution for the extrinsic parameters \citep{2016arXiv160904470M}.
\color{black}

@@

Until now we have considered $k_0$ and $k_1$ without association with any specific dust model, but they could be interpreted as
parameters of a single  dust model.
\color{red}
Our model for broadband photometry does not specify how spectral features and color parameters affect the supernova spectrum, information
necessary to predict the magnitude extinction of dust models.
\color{blue}
The lack of correspondence between dust models and broadband extinction is quantified when we apply an $R^F_V=2.5$ SN-frame dust to the
observed spectra used in our sample, to yield a distribution in the quantity $R_V=A_V/(A_B-A_V)$ with mean and standard deviation of $3.11 \pm 0.35$.
While an accurate study of fundamental dust properties would incorporate the dust law within the model itself, we now proceed nonetheless
in making a connection between our model outputs and dust.

\color{red}
Plotted in Figure~\ref{avebv:fig} are the expectations for measured $A(V)$, $E(B-V)$, and $R^T_V$ for $R^F_V=1.1, 1.9, 2.5, 3.1$
within the context of the \citetalias{1999PASP..111...63F} model.
Note that for our broadband measurements, the relationship between $A(V)$ and $E(B-V)$ is subtly
non-linear, and $R^T_V$ is not constant for a fixed value of $R^F_V$.
\color{black}

The \citet{1989ApJ...345..245C} dust law can be expressed as a two-parameter linear model
\begin{equation}
A_X = a(X)  A_V + b(X) E(B-V).
\end{equation}
The \citetalias{1999PASP..111...63F} model is approximately described by a linear model with
the above form, with
$a(U,B,V,R,I)=\{0.96,   1.00,   1.00,   0.97,   0.78\}$ and $b(U,B,V,R,I)=\{  1.64,   1.06,   0.16,  -0.40,  -0.52\}$ for the case of
$R^F_V=2.5$
\color{red}
and $A^F_V=0.1$,
for the measurement of the updated
\citet{2007ApJ...663.1187H} SN~Ia template at $B$-band peak as seen through our synthetic passbands.
\color{black}
This linear approximation
is good in the ranges of
\color{red}
 $R^F_V$ and $A^F_V$,
\color{black}
and wavelengths under consideration, in that the values of the components  vary by $<5$\%.

@@

\section{Supernova Model II: Two Extrinsic Parameters, One Intrinsic Parameter}
\label{model2:sec}
In \S\ref{model:sec} we show that the SNfactory data analyzed with an agnostic model with
two extrinsic parameters produce
magnitudes consistent
with two-parameter dust models that have been established for stars in the Milky Way, LMC, and SMC.
An extension to our model is now added to explore the possibility of
a possible additional
color parameter that could have its origins in supernova astrophysics.
The extended model is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_0+\alpha_0 EW_{Ca} + \beta_0 EW_{Si} + \eta_0 \lambda_{Si} + \delta_0 D\\
c_1+\alpha_1 EW_{Ca} + \beta_1 EW_{Si} + \eta_1 \lambda_{Si} + \delta_1 D \\
c_2+\alpha_2 EW_{Ca} + \beta_2 EW_{Si} + \eta_2 \lambda_{Si} + \delta_2 D\\
c_3+\alpha_3 EW_{Ca} + \beta_3 EW_{Si} + \eta_3 \lambda_{Si} + \delta_3 D\\
c_4+\alpha_4 EW_{Ca} + \beta_4 EW_{Si}+ \eta_4 \lambda_{Si} + \delta_4 D
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{0} k_0 +\gamma^1_{0} k_1 \\B +\gamma^0_{1} k_0 +\gamma^1_{1} k_1 \\
V+\gamma^0_{2} k_0+\gamma^1_{2} k_1\\R+\gamma^0_{3} k_0 + \gamma^1_{3} k_1\\I+\gamma^0_{4} k_0+\gamma^1_{4} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}
In all aspects it is identical to the previous model except for the addition of the term $\delta D$,
which affects the mean
absolute magnitudes:
each supernova has an intrinsic $D$ parameter that is linearly related to
absolute magnitudes through global coefficients $\delta$.  As with $k_0$ and $k_1$,  $D$ is a latent supernova parameter whose existence is inferred from
its effects on broadband magnitudes.
